# Homework 7

```{r}
library(ISLR)
library(rms)
library(car)

data(Wage)
dd_wage <- datadist(Wage)
options(datadist = "dd_wage")
```

# Question 1
```{r}
fit_lin_lm <- lm(wage ~ age + education + race, data = Wage)
fit_lin_rms <- ols(wage ~ age + education + race, data = Wage)

crPlots(fit_lin_lm, terms = ~ age, 
        main = "component and residual plot",
        col = "blue", pch = 16)

plot(rms::Predict(fit_lin_rms, age), 
     xlab = "age", ylab = "predicted wage",
     main = "pep")
```

Based on the age and wage relationship, there is a slight non linear relationship.

# Question 2

```{r}
fit_spline <- ols(wage ~ rcs(age, 4) + education + race, data = Wage, x = TRUE, y = TRUE)

plot(rms::Predict(fit_spline, age), 
     xlab = "age", ylab = "predicted wage",
     main = "pep")
```

There is a large difference in the implied curve compared to the linear fit. The implied curve has a large increase until around age 40, and then a gradual decrease.

# Question 3
```{r}
set.seed(7)
k <- 10
folds <- sample(1:k, nrow(Wage), replace = TRUE)

mse_lin <- numeric(k)
mse_spline <- numeric(k)

for(i in 1:k) {
  train_data <- Wage[folds != i, ]
  test_data  <- Wage[folds == i, ]
  
  m_lin <- lm(wage ~ age + education + race, data = train_data)
  m_spl <- lm(wage ~ rcs(age, 4) + education + race, data = train_data)
  
  pred_lin <- predict(m_lin, newdata = test_data)
  pred_spl <- predict(m_spl, newdata = test_data)
  
  mse_lin[i] <- mean((test_data$wage - pred_lin)^2)
  mse_spline[i] <- mean((test_data$wage - pred_spl)^2)
}

cat("linear model:", mean(mse_lin), "\n")
cat("spline model:", mean(mse_spline), "\n")
```

Yes, based on the mse values, the spline model has a lower mean squared error than the linear model.

# Question 4
```{r}
fit_lin_cal <- ols(wage ~ age + education + race, data = Wage, x = TRUE, y = TRUE)

cal_lin <- calibrate(fit_lin_cal, method = "boot", B = 100)
cal_spline <- calibrate(fit_spline, method = "boot", B = 100)

par(mfrow = c(1, 2))

plot(cal_lin, main = "linear", 
     xlab = "predicted wage", ylab = "observed wage")

plot(cal_spline, main = "spline", 
     xlab = "predicted wage", ylab = "observed wage")

par(mfrow = c(1, 1))
```

The spline model seems to have slight deviation at it's early values and therfore slight systematic miscalibration.

# Question 5

Based on the following:
- Our initial answer to question 1 showed that there was non linear relationship.
- The spline model in question 3 had a lower mean squared error than the linear model.
- Although the spline model in question 4 had slight miscalibration, it is still well calibrated in terms of fitting the 45 degree line.

