# Homework 8

```{r}
library(glmnet)
library(pls)
riboflavin <- readRDS("riboflavin.rds")

X <- riboflavin$x
y <- riboflavin$y 
```

# Question 1
```{r}
d_full <- data.frame(y = y, x = X)
fit_ols <- lm(y ~ ., data = d_full)

sum(is.na(coef(fit_ols)))
```
We get 4018 NA coefficients. This is because there is a smaller sample size compared to the number of predictors. This is found as n-1 coefficients so the remaining coefficients are left as NA after calucating this formula.

# Question 2

```{r}
set.seed(7)

cv_ridge <- cv.glmnet(X, y, alpha = 0, nfolds = 10)

cv_ridge$lambda.min
cv_ridge$lambda.1se
plot(cv_ridge)
```

As lambda increases, the mean squared error decreases on our log converted plot.

# Question 3
```{r}
fit_ridge <- glmnet(X, y, alpha = 0)

plot(fit_ridge, xvar = "lambda", label = FALSE, 
     main = "Ridge Coefficient Path")
```

Based on our plot, the coefficients start at various values and slowly converge to 0 as lambda increase. Based on the number of coefficients, we can see that all 4088 coefficients are still retained. This means that no coefficient actually reaches 0.

# Question 4
```{r}
set.seed(7)
cv_lasso <- cv.glmnet(X, y, alpha = 1, nfolds = 10)

lasso_coefs <- coef(cv_lasso, s = "lambda.1se")
features_retained <- sum(lasso_coefs != 0) - 1 
print(features_retained)

par(mfrow = c(1, 2))
plot(cv_lasso, main = "Lasso CV Error")
fit_lasso <- glmnet(X, y, alpha = 1)
plot(fit_lasso, xvar = "lambda", main = "Lasso Path")
par(mfrow = c(1, 1))
```

# Question 5

```{r}
set.seed(7)
alphas <- c(0.25, 0.5, 0.75)

en_results <- data.frame(
  Alpha = numeric(),
  Lambda_min = numeric(),
  CV_MSE = numeric(),
  Features_Selected = numeric()
)

for (a in alphas) {
  set.seed(7) 
  cv_en <- cv.glmnet(X, y, alpha = a, nfolds = 10)
  
  best_mse <- min(cv_en$cvm)
  best_lambda <- cv_en$lambda.min
  
  en_coefs <- coef(cv_en, s = "lambda.min")
  features <- sum(en_coefs != 0) - 1
  en_results <- rbind(en_results, data.frame(
    Alpha = a,
    Lambda_min = best_lambda,
    CV_MSE = best_mse,
    Features_Selected = features
  ))
}

print(en_results)
```

Alpha 0.25 shows the lowest CV-MSE. There are generally more features for all values of alpha compared to the lasso.

# Question 6

```{r}
set.seed(7)
fit_pcr <- pcr(y ~ x, data = riboflavin, scale = TRUE, validation = "CV")

validationplot(fit_pcr, val.type = "MSE", main = "PCR: CV MSE")
msep_vals <- MSEP(fit_pcr)$val[1, 1, ]
k_opt <- which.min(msep_vals) - 1
print(k_opt)
beta_pcr <- coef(fit_pcr, ncomp = k_opt)

pred_1 <- predict(fit_pcr, newdata = riboflavin$x[1, , drop=FALSE], ncomp = k_opt)
print(pred_1)
```

Because each k principal component is used to build the model for the 4088 coeffecients, each gene expression value must be measured/used to calculate the new model.

# Question 7

```{r}
c(
  Ridge = min(cv_ridge$cvm),
  Lasso = min(cv_lasso$cvm),
  ElasticNet = min(cv_en$cvm),
  PCR = min(msep_vals)
)
```

The lasso method achieves the lowest CV-MSE on the riboflavin dataset.

Based on the results, the lasso method works better here. Not only is the CV-MSE lower than the ridge method, it also has fewer selected features that makes it focus on predictors that change the outcome more meaningfully.